---
title: "R-Based Forecasting Project: Construction Cost Index (2015=100)"
author: "Adın Soyadın - Öğrenci Numarası"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 1. Project Overview

This project forecasts the Construction Cost Index (2015=100) published by TurkStat (TÜİK) using monthly time series data accessed directly through the `tuikr` R package. The forecast target is February 2026, the next period after the latest available observation (January 2026).

## 2. Data Access via tuikr

```{r data-access}
library(tuikr)
library(httr)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

tables_price <- statistical_tables(theme = "6")

url_cci <- tables_price$table_url[
  tables_price$table_name == "Construction Cost Index (2015=100) and Rate of Change "]

response_cci <- GET(
  url_cci,
  add_headers(
    "User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36",
    "Accept"     = "application/json, text/plain, */*",
    "Referer"    = "https://veriportali.tuik.gov.tr/"
  )
)

cat("HTTP Status:", status_code(response_cci), "\n")
cat("Data access date:", format(Sys.time(), "%Y-%m-%d"), "\n")

raw_cci <- content(response_cci, as = "raw")
tmp_cci <- tempfile(fileext = ".xls")
writeBin(raw_cci, tmp_cci)
df_cci  <- read_excel(tmp_cci)
```

## 3. Data Preparation

```{r data-prep}
colnames(df_cci) <- c("year","Jan","Feb","Mar","Apr","May","Jun",
                       "Jul","Aug","Sep","Oct","Nov","Dec")

df_index <- df_cci[6:17, ]

df_long <- df_index %>%
  mutate(year = as.integer(year)) %>%
  pivot_longer(cols = Jan:Dec, names_to = "month", values_to = "index") %>%
  mutate(
    month_num = match(month, c("Jan","Feb","Mar","Apr","May","Jun",
                                "Jul","Aug","Sep","Oct","Nov","Dec")),
    index = as.numeric(index)
  ) %>%
  filter(!is.na(index)) %>%
  arrange(year, month_num)

ts_cci <- ts(df_long$index, start = c(2015, 1), frequency = 12)

cat("Latest available observation: 2026 January\n")
cat("Forecast target period: 2026 February\n")
cat("Number of observations:", nrow(df_long), "\n")
```

## 4. Exploratory Time Series Analysis

```{r eda-plot}
plot_data <- data.frame(
  date  = as.numeric(time(ts_cci)),
  index = as.numeric(ts_cci)
)

ggplot(plot_data, aes(x = date, y = index)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 1.5) +
  labs(
    title    = "Construction Cost Index (2015=100)",
    subtitle = "Turkey, Monthly (January 2015 - January 2026)",
    x = "Year", y = "Index Value",
    caption  = "Source: TurkStat (TÜİK)"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

The series shows a strong upward trend, especially from 2021 onward due to high inflation in Turkey. Seasonality is weak relative to the dominant trend. No missing values detected.

## 5. Naive Forecasting

```{r naive}
naive_forecast <- as.numeric(tail(ts_cci, 1))
naive_fitted   <- c(NA, as.numeric(ts_cci[-length(ts_cci)]))
naive_errors   <- as.numeric(ts_cci) - naive_fitted

naive_bias <- mean(naive_errors, na.rm = TRUE)
naive_mad  <- mean(abs(naive_errors), na.rm = TRUE)
naive_mse  <- mean(naive_errors^2, na.rm = TRUE)
naive_mape <- mean(abs(naive_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
naive_rsfe <- sum(naive_errors, na.rm = TRUE)
naive_ts   <- naive_rsfe / naive_mad

cat("Naive Forecast Feb 2026:", round(naive_forecast, 2), "\n")
cat("Bias:", round(naive_bias, 2), "| MAD:", round(naive_mad, 2),
    "| MSE:", round(naive_mse, 2), "| MAPE:", round(naive_mape, 2), "%\n")
cat("RSFE:", round(naive_rsfe, 2), "| Tracking Signal:", round(naive_ts, 2), "\n")

naive_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = naive_fitted
)

ggplot(naive_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Naive Forecast"),
            linewidth = 0.8, linetype = "dashed", na.rm = TRUE) +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Naive Forecast" = "red")) +
  labs(title = "Actual vs Naive Forecast", x = "Year",
       y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 6. Moving Average

```{r moving-average}
window     <- 3
ma_fitted  <- as.numeric(stats::filter(ts_cci, rep(1/window, window), sides = 1))
ma_forecast <- mean(tail(as.numeric(ts_cci), window))

ma_errors <- as.numeric(ts_cci) - ma_fitted
ma_bias <- mean(ma_errors, na.rm = TRUE)
ma_mad  <- mean(abs(ma_errors), na.rm = TRUE)
ma_mse  <- mean(ma_errors^2, na.rm = TRUE)
ma_mape <- mean(abs(ma_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
ma_rsfe <- sum(ma_errors, na.rm = TRUE)
ma_ts   <- ma_rsfe / ma_mad

cat("MA(3) Forecast Feb 2026:", round(ma_forecast, 2), "\n")
cat("Bias:", round(ma_bias, 2), "| MAD:", round(ma_mad, 2),
    "| MSE:", round(ma_mse, 2), "| MAPE:", round(ma_mape, 2), "%\n")
cat("RSFE:", round(ma_rsfe, 2), "| Tracking Signal:", round(ma_ts, 2), "\n")

ma_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = ma_fitted
)

ggplot(ma_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "MA(3)"),
            linewidth = 0.8, linetype = "dashed", na.rm = TRUE) +
  scale_color_manual(values = c("Actual" = "steelblue", "MA(3)" = "darkgreen")) +
  labs(title = "Actual vs Moving Average (3-month)", x = "Year",
       y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 7. Weighted Moving Average

```{r wma}
w <- c(0.1, 0.3, 0.6)
n <- length(ts_cci)
wma_fitted <- rep(NA, n)
for (i in 3:n) {
  wma_fitted[i] <- sum(w * as.numeric(ts_cci)[(i-2):i])
}
wma_forecast <- sum(w * tail(as.numeric(ts_cci), 3))

wma_errors <- as.numeric(ts_cci) - wma_fitted
wma_bias <- mean(wma_errors, na.rm = TRUE)
wma_mad  <- mean(abs(wma_errors), na.rm = TRUE)
wma_mse  <- mean(wma_errors^2, na.rm = TRUE)
wma_mape <- mean(abs(wma_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
wma_rsfe <- sum(wma_errors, na.rm = TRUE)
wma_ts   <- wma_rsfe / wma_mad

cat("WMA Forecast Feb 2026:", round(wma_forecast, 2), "\n")
cat("Bias:", round(wma_bias, 2), "| MAD:", round(wma_mad, 2),
    "| MSE:", round(wma_mse, 2), "| MAPE:", round(wma_mape, 2), "%\n")
cat("RSFE:", round(wma_rsfe, 2), "| Tracking Signal:", round(wma_ts, 2), "\n")

wma_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = wma_fitted
)

ggplot(wma_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "WMA"),
            linewidth = 0.8, linetype = "dashed", na.rm = TRUE) +
  scale_color_manual(values = c("Actual" = "steelblue", "WMA" = "orange")) +
  labs(title = "Actual vs Weighted Moving Average (0.1, 0.3, 0.6)",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 8. Exponential Smoothing

```{r exp_smoothing}
library(forecast)

# Test alpha = 0.9 (high - suitable for trending series)
es_model <- ses(ts_cci, alpha = 0.9, h = 1)
es_forecast <- as.numeric(es_model$mean)
es_fitted <- as.numeric(es_model$fitted)

es_errors <- as.numeric(ts_cci) - es_fitted
es_bias <- mean(es_errors, na.rm = TRUE)
es_mad  <- mean(abs(es_errors), na.rm = TRUE)
es_mse  <- mean(es_errors^2, na.rm = TRUE)
es_mape <- mean(abs(es_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
es_rsfe <- sum(es_errors, na.rm = TRUE)
es_ts   <- es_rsfe / es_mad

cat("Exponential Smoothing (alpha=0.9) Forecast Feb 2026:", 
    round(es_forecast, 2), "\n")
cat("Bias:", round(es_bias, 2), "| MAD:", round(es_mad, 2),
    "| MSE:", round(es_mse, 2), "| MAPE:", round(es_mape, 2), "%\n")
cat("RSFE:", round(es_rsfe, 2), "| Tracking Signal:", round(es_ts, 2), "\n")

es_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = es_fitted
)

ggplot(es_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "ES (alpha=0.9)"),
            linewidth = 0.8, linetype = "dashed") +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "ES (alpha=0.9)" = "purple")) +
  labs(title = "Actual vs Exponential Smoothing (alpha=0.9)",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 9. Trend-Adjusted Exponential Smoothing (Holt's Method)

```{r trend-es}
# Holt's method - suitable since series has strong trend
holt_model <- holt(ts_cci, alpha = 0.9, beta = 0.3, h = 1)
holt_forecast <- as.numeric(holt_model$mean)
holt_fitted <- as.numeric(holt_model$fitted)

holt_errors <- as.numeric(ts_cci) - holt_fitted
holt_bias <- mean(holt_errors, na.rm = TRUE)
holt_mad  <- mean(abs(holt_errors), na.rm = TRUE)
holt_mse  <- mean(holt_errors^2, na.rm = TRUE)
holt_mape <- mean(abs(holt_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
holt_rsfe <- sum(holt_errors, na.rm = TRUE)
holt_ts   <- holt_rsfe / holt_mad

cat("Holt's Forecast Feb 2026:", round(holt_forecast, 2), "\n")
cat("Bias:", round(holt_bias, 2), "| MAD:", round(holt_mad, 2),
    "| MSE:", round(holt_mse, 2), "| MAPE:", round(holt_mape, 2), "%\n")
cat("RSFE:", round(holt_rsfe, 2), "| Tracking Signal:", round(holt_ts, 2), "\n")

holt_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = holt_fitted
)

ggplot(holt_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Holt's Method"),
            linewidth = 0.8, linetype = "dashed") +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Holt's Method" = "darkred")) +
  labs(title = "Actual vs Trend-Adjusted Exponential Smoothing (Holt's)",
       subtitle = "alpha=0.9, beta=0.3",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 10. Linear Trend Projection

```{r linear-trend}
# Create time index
t <- 1:length(ts_cci)
trend_model <- lm(as.numeric(ts_cci) ~ t)
summary(trend_model)

trend_fitted <- as.numeric(fitted(trend_model))
trend_forecast <- predict(trend_model, newdata = data.frame(t = length(ts_cci) + 1))

trend_errors <- as.numeric(ts_cci) - trend_fitted
trend_bias <- mean(trend_errors)
trend_mad  <- mean(abs(trend_errors))
trend_mse  <- mean(trend_errors^2)
trend_mape <- mean(abs(trend_errors / as.numeric(ts_cci)) * 100)
trend_rsfe <- sum(trend_errors)
trend_ts   <- trend_rsfe / trend_mad

cat("Linear Trend Forecast Feb 2026:", round(trend_forecast, 2), "\n")
cat("Intercept:", round(coef(trend_model)[1], 2), 
    "| Slope:", round(coef(trend_model)[2], 2), "\n")
cat("Bias:", round(trend_bias, 2), "| MAD:", round(trend_mad, 2),
    "| MSE:", round(trend_mse, 2), "| MAPE:", round(trend_mape, 2), "%\n")
cat("RSFE:", round(trend_rsfe, 2), "| Tracking Signal:", round(trend_ts, 2), "\n")

trend_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = trend_fitted
)

ggplot(trend_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Linear Trend"),
            linewidth = 0.8, linetype = "dashed") +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Linear Trend" = "brown")) +
  labs(title = "Actual vs Linear Trend Projection",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```
## 11. Seasonal Indices

```{r seasonal_indices}
# Monthly seasonal indices
# Extract month from time series
months <- cycle(ts_cci)
years  <- floor(time(ts_cci))

# Calculate seasonal indices
monthly_means <- tapply(as.numeric(ts_cci), months, mean)
overall_mean  <- mean(as.numeric(ts_cci))
seasonal_idx  <- monthly_means / overall_mean

cat("Seasonal Indices by Month:\n")
month_names <- c("Jan","Feb","Mar","Apr","May","Jun",
                  "Jul","Aug","Sep","Oct","Nov","Dec")
for(i in 1:12) {
  cat(month_names[i], ":", round(seasonal_idx[i], 4), "\n")
}

# The series has a dominant trend; seasonal variation is minimal
# Deseasonalized values
deseasonalized <- as.numeric(ts_cci) / seasonal_idx[months]

# Seasonal indices plot
si_df <- data.frame(
  month = month_names,
  index = as.numeric(seasonal_idx)
)
si_df$month <- factor(si_df$month, levels = month_names)

ggplot(si_df, aes(x = month, y = index)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  labs(title = "Seasonal Indices by Month",
       subtitle = "Values above 1 indicate above-average months",
       x = "Month", y = "Seasonal Index") +
  theme_minimal()
```

## 12. Additive Decomposition

```{r additive_decomp}
# Additive decomposition
decomp_add <- decompose(ts_cci, type = "additive")

# Plot decomposition
plot(decomp_add)

# Forecast using trend + seasonal
trend_comp    <- as.numeric(decomp_add$trend)
seasonal_comp <- as.numeric(decomp_add$seasonal)
random_comp   <- as.numeric(decomp_add$random)

# Fitted values
add_fitted <- trend_comp + seasonal_comp
add_errors <- as.numeric(ts_cci) - add_fitted
add_bias <- mean(add_errors, na.rm = TRUE)
add_mad  <- mean(abs(add_errors), na.rm = TRUE)
add_mse  <- mean(add_errors^2, na.rm = TRUE)
add_mape <- mean(abs(add_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
add_rsfe <- sum(add_errors, na.rm = TRUE)
add_ts   <- add_rsfe / add_mad

cat("Additive Decomposition:\n")
cat("Bias:", round(add_bias, 2), "| MAD:", round(add_mad, 2),
    "| MSE:", round(add_mse, 2), "| MAPE:", round(add_mape, 2), "%\n")
cat("RSFE:", round(add_rsfe, 2), "| Tracking Signal:", round(add_ts, 2), "\n")

add_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = add_fitted
)

ggplot(add_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Additive Decomp"),
            linewidth = 0.8, linetype = "dashed", na.rm = TRUE) +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Additive Decomp" = "darkgreen")) +
  labs(title = "Actual vs Additive Decomposition",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 13. Multiplicative Decomposition

```{r mult_decomp}
decomp_mult <- decompose(ts_cci, type = "multiplicative")

plot(decomp_mult)

trend_m    <- as.numeric(decomp_mult$trend)
seasonal_m <- as.numeric(decomp_mult$seasonal)

mult_fitted <- trend_m * seasonal_m
mult_errors <- as.numeric(ts_cci) - mult_fitted
mult_bias <- mean(mult_errors, na.rm = TRUE)
mult_mad  <- mean(abs(mult_errors), na.rm = TRUE)
mult_mse  <- mean(mult_errors^2, na.rm = TRUE)
mult_mape <- mean(abs(mult_errors / as.numeric(ts_cci)) * 100, na.rm = TRUE)
mult_rsfe <- sum(mult_errors, na.rm = TRUE)
mult_ts   <- mult_rsfe / mult_mad

cat("Multiplicative Decomposition:\n")
cat("Bias:", round(mult_bias, 2), "| MAD:", round(mult_mad, 2),
    "| MSE:", round(mult_mse, 2), "| MAPE:", round(mult_mape, 2), "%\n")
cat("RSFE:", round(mult_rsfe, 2), "| Tracking Signal:", round(mult_ts, 2), "\n")

mult_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = mult_fitted
)

ggplot(mult_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Multiplicative Decomp"),
            linewidth = 0.8, linetype = "dashed", na.rm = TRUE) +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Multiplicative Decomp" = "darkred")) +
  labs(title = "Actual vs Multiplicative Decomposition",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 14. Regression with Trend and Seasonal Dummies

```{r regression_seasonal}
n <- length(ts_cci)
t <- 1:n
months_factor <- as.factor(cycle(ts_cci))

reg_model <- lm(as.numeric(ts_cci) ~ t + months_factor)
summary(reg_model)

reg_fitted   <- as.numeric(fitted(reg_model))
reg_forecast <- predict(reg_model, 
                         newdata = data.frame(t = n + 1, 
                                              months_factor = factor(2, 
                                              levels = levels(months_factor))))

reg_errors <- as.numeric(ts_cci) - reg_fitted
reg_bias <- mean(reg_errors)
reg_mad  <- mean(abs(reg_errors))
reg_mse  <- mean(reg_errors^2)
reg_mape <- mean(abs(reg_errors / as.numeric(ts_cci)) * 100)
reg_rsfe <- sum(reg_errors)
reg_ts   <- reg_rsfe / reg_mad

cat("Regression with Trend+Seasonal Forecast Feb 2026:", 
    round(reg_forecast, 2), "\n")
cat("Bias:", round(reg_bias, 2), "| MAD:", round(reg_mad, 2),
    "| MSE:", round(reg_mse, 2), "| MAPE:", round(reg_mape, 2), "%\n")
cat("RSFE:", round(reg_rsfe, 2), "| Tracking Signal:", round(reg_ts, 2), "\n")

reg_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = reg_fitted
)

ggplot(reg_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Regression"),
            linewidth = 0.8, linetype = "dashed") +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Regression" = "darkviolet")) +
  labs(title = "Actual vs Regression with Trend and Seasonal Dummies",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```
## 15. Forecast Accuracy Comparison

```{r comparison_table}
library(knitr)

comparison <- data.frame(
  Method = c("Naive", "Moving Average (3)", "Weighted Moving Average",
             "Exponential Smoothing", "Trend-Adjusted ES (Holt)",
             "Linear Trend", "Seasonal Indices", 
             "Additive Decomposition", "Multiplicative Decomposition",
             "Regression Trend+Seasonal"),
  Bias = round(c(naive_bias, ma_bias, wma_bias, es_bias, holt_bias,
                  trend_bias, NA, add_bias, mult_bias, reg_bias), 2),
  MAD  = round(c(naive_mad, ma_mad, wma_mad, es_mad, holt_mad,
                  trend_mad, NA, add_mad, mult_mad, reg_mad), 2),
  MSE  = round(c(naive_mse, ma_mse, wma_mse, es_mse, holt_mse,
                  trend_mse, NA, add_mse, mult_mse, reg_mse), 2),
  MAPE = round(c(naive_mape, ma_mape, wma_mape, es_mape, holt_mape,
                  trend_mape, NA, add_mape, mult_mape, reg_mape), 2),
  RSFE = round(c(naive_rsfe, ma_rsfe, wma_rsfe, es_rsfe, holt_rsfe,
                  trend_rsfe, NA, add_rsfe, mult_rsfe, reg_rsfe), 2),
  Tracking_Signal = round(c(naive_ts, ma_ts, wma_ts, es_ts, holt_ts,
                              trend_ts, NA, add_ts, mult_ts, reg_ts), 2),
  Next_Period_Forecast = round(c(naive_forecast, ma_forecast, wma_forecast,
                                  es_forecast, holt_forecast, trend_forecast,
                                  NA, NA, NA, reg_forecast), 2)
)

kable(comparison, caption = "Forecast Accuracy Comparison Table")
```

## 16. Superior Method Selection

```{r superior_method}
cat("=== SUPERIOR METHOD SELECTION ===\n\n")
cat("Selected Method: Trend-Adjusted Exponential Smoothing (Holt's Method)\n\n")
cat("Justification:\n")
cat("1. The CCI series shows a dominant upward trend throughout 2015-2026.\n")
cat("2. Holt's method explicitly models both level and trend components.\n")
cat("3. Alpha=0.9 gives high weight to recent observations (appropriate for\n")
cat("   rapidly changing inflation-driven index).\n")
cat("4. Beta=0.3 allows moderate trend adaptation.\n")
cat("5. MAPE is competitive and tracking signal is acceptable.\n")
cat("6. Unlike linear trend, Holt's captures non-linear acceleration.\n\n")
cat("Forecast for February 2026:", round(holt_forecast, 2), "\n")

# Final plot
final_df <- data.frame(
  date   = as.numeric(time(ts_cci)),
  actual = as.numeric(ts_cci),
  fitted = holt_fitted
)

# Add forecast point
forecast_point <- data.frame(
  date   = 2026 + 1/12,
  actual = NA,
  fitted = holt_forecast
)

ggplot(final_df, aes(x = date)) +
  geom_line(aes(y = actual, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = fitted, color = "Holt's ES"), 
            linewidth = 0.8, linetype = "dashed") +
  geom_point(data = forecast_point, 
             aes(x = date, y = fitted), 
             color = "darkred", size = 4, shape = 17) +
  annotate("text", x = 2026.1, y = holt_forecast + 50,
           label = paste0("Feb 2026\nForecast: ", round(holt_forecast, 2)),
           color = "darkred", size = 3.5) +
  scale_color_manual(values = c("Actual" = "steelblue",
                                 "Holt's ES" = "darkred")) +
  labs(title = "Superior Method: Trend-Adjusted Exponential Smoothing",
       subtitle = "Final Forecast for February 2026",
       x = "Year", y = "Index Value", color = "") +
  theme_minimal() +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2026)
```

## 17. Final Forecast Summary

```{r final_forecast}
final_result <- data.frame(
  Item  = c("Selected Superior Method", "Data Access Date",
             "Latest Available Observation", "Forecast Target Period",
             "Forecasted Value", "Alpha", "Beta"),
  Value = c("Trend-Adjusted Exponential Smoothing (Holt's Method)",
             format(Sys.Date(), "%Y-%m-%d"),
             "January 2026", "February 2026",
             round(holt_forecast, 2), "0.9", "0.3")
)

kable(final_result, caption = "Final Forecast Result")

cat("\nInterpretation:\n")
cat("The Construction Cost Index is forecast to reach", 
    round(holt_forecast, 2), "in February 2026.\n")
cat("This represents a continued upward trend driven by inflation\n")
cat("and rising construction material costs in Turkey.\n")
```

## 18. Limitations

The following limitations should be considered when interpreting this forecast:

- The series contains a structural break around 2021-2022 due to Turkey's high inflation period, which may reduce forecast accuracy.
- Seasonal patterns are weak compared to the trend, so seasonal methods add limited value.
- Only 133 monthly observations are available (2015-2026), which limits the robustness of decomposition methods.
- The forecast is for one period ahead only; longer-horizon forecasts would carry much higher uncertainty.
- External shocks (policy changes, commodity price spikes) cannot be captured by time series methods alone.

## 19. Reproducibility

This project can be reproduced by:

1. Installing R 4.6.0 and RStudio
2. Installing the `tuikr` package from GitHub: `pak::pak("emraher/tuikr")`
3. Installing required packages: `httr`, `readxl`, `dplyr`, `tidyr`, `ggplot2`, `scales`, `forecast`
4. Opening `forecasting_project.Rmd` and clicking **Knit**
5. Data is accessed directly from TÜİK via `tuikr::statistical_tables()` — no manual download required